---
title: "Trait responses to global change"
format: 
  html:
    self-contained: true
editor: visual
execute:
  warnings: false
  messages: false
---

```{r}
#| label: set-up
#| echo: false
#| message: false

library(targets)
library(tarchetypes)
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
library(gt)
library(gtExtras)
library(broom)

```

# Results

### Plant functional traits

```{r}
#| label: fig-nitrogen
#| echo: false
#| warning: false
#| fig-height: 6
#| fig-cap: Plant community functional trait responses to warming and nitrogen addition. Bootstrapped trait means increasing nitrogen addition (log-transformed) under ambient (grey) and warm (pink) climate originating from alpine (triangle, solid line) and subalpine (circle, dashed line) sites. The points show the raw data and the lines, the predictions and confidence intervals from the model. The letters in the bottom right corner indicate significant effects of W = warming, N = nitrogen addition, and O = origin community, with WxO and NxO representing their interaction. 

tar_read(nitrogen_figure)
```

```{r}
#| label: fig-grazing
#| echo: false
#| warning: false
#| fig-height: 6
#| fig-cap: Plant community functional trait responses to warming and grazing. Bootstrapped trait means under a) ambient (grey) and warm climate (pink), b) grazed (blue) and ungrazed (orange) conditions, originating from subalpine and alpine sites. The points show the raw data and the lines, the predictions and confidence intervals from the model. The letters in the bottom right corner indicate significant effects of W = warming, G = grazing, and O = origin community, with WxO and GxO representing their interaction. 

tar_read(grazing_figure)
```

```{r}
#| label: tbl-single-traits
#| echo: false
#| eval: false
#| tbl-cap: The main and interactive linear model effects for the warming by nitrogen and warming by grazing models. Each model included effects for warming (W), origin (O), model-specific drivers—nitrogen (N) and grazing (G)—and an interaction term for pair-wise interactions. Significant effects (p < 0.05) are bolded.

tar_load(trait_anovas)

df <- trait_anovas |> 
    mutate(Traits = case_when(Traits == "Plant~height~(cm)" ~ "Plant height (cm)",
                              Traits == "Leaf~dry~mass~(g)" ~ "Leaf dry mass (cm)",
                              Traits == "Leaf~area~(cm^2)" ~ "Leaf area (cm<sup>2</sup>)",
                              Traits == "Leaf~thickness~(mm)" ~ "Leaf thickness (mm)",
                              Traits == "SLA~(cm^2*g^{-1})" ~ "SLA (cm<sup>2</sup>g<sup>-1</sup>)",
                              TRUE ~ Traits),
           Traits = purrr::map(Traits, gt::html)) |> 
  group_by(Traits) |> 
  gt() |> 
  tab_spanner(
      label = md('Warming x nitrogen'),
      columns = c('Term', 'Sum of Square', 'df', 'F', 'P')
    ) |> 
  tab_spanner(
      label = md('Warming x grazing'),
      columns = c('Term_g', 'Sum of Square_g', 'df_g', 'F_g', 'P_g')
    ) |> 
  tab_style(
        style = cell_borders(sides = "left", color = "white", weight = px(15), style = "solid"),
        locations = cells_body(
          columns = "Term_g")
      ) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c("Term", 'Sum of Square', df, 'F', P),
      rows = P <= 0.05)
    )|> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c("Term_g", 'Sum of Square_g', df_g, F_g, P_g),
      rows = P_g <= 0.05)
    )|> 
  cols_label(
    Term_g = "Term",
    'Sum of Square_g' = "Sum of Square",
    df_g = "df",
    F_g = "F",
    P_g = "P"
  ) |> 
  tab_options(
  table.font.size = 10,
  data_row.padding = gt::px(1)
  )

df
#df %>% gtsave(filename = "Table_1.png")
```

```{r}
#| label: single-traits_stats
#| echo: false

tar_load(trait_prediction)

df <- trait_prediction |>
  ungroup() |>
  select(figure_names, anova_tidy) |> 
  unnest(anova_tidy) |> 
  mutate(sumsq = round(sumsq, 2),
               df = round(df, 2),
               statistic = round(statistic, 2),
               p.value = round(p.value, 3)) |>
  mutate(term = case_when(str_detect(term, "warming:Nitrogen_log") ~ "WxN",
                            str_detect(term, "Nitrogen_log:origSiteID") ~ "NxO",
                            str_detect(term, "warming:origSiteID") ~ "WxO",
                            str_detect(term, "warming:grazing") ~ "WxG",
                            str_detect(term, "origSiteID:grazing") ~ "GxO",
                            str_detect(term, "origSiteID") ~ "O",
                            str_detect(term, "Nitrogen_log") ~ "N",
                            str_detect(term, "warming") ~ "W",
                            str_detect(term, "grazing") ~ "G",
                            TRUE ~ term)) |>
  rename(Traits = figure_names, Term = term, "Sum of Square" = sumsq, "F" = statistic, "P" = "p.value") |> 
  select(Traits, Term, `Sum of Square`, df, 'F', P) |>
  arrange(Traits)

# df2 <- df |>
#   group_by(Traits) |>
#   gt() |>
#   tab_style(
#         style = cell_borders(sides = "left", color = "white", weight = px(15), style = "solid"),
#         locations = cells_body(
#           columns = "Term")
#       ) |>
#   tab_style(
#     style = list(
#       cell_text(weight = "bold")
#       ),
#     locations = cells_body(
#       columns = c("Term", 'Sum of Square', df, 'F', P),
#       rows = P <= 0.05)
#     )|>
#   tab_options(
#   table.font.size = 10,
#   data_row.padding = gt::px(1)
#   )
# 
# df2
#df2 %>% gtsave(filename = "Table_1.png")

```

```{r}
#| label: tbl-label
#| echo: false
#| layout-ncol: 2
#| tbl-cap: The main and interactive linear model effects for the warming by nitrogen and warming by grazing models. Each model included effects for warming (W), origin community (O), model-specific drivers—nitrogen (N) and grazing (G)—and an interaction term for pair-wise interactions. Significant effects (p < 0.05) are bolded. SLA is specific leaf area and LDMC is leaf dry matter content.


library(knitr)

# table on the left
df |> 
  slice(1:30) |> 
  group_by(Traits) |> 
  gt() |> 
  text_transform(
    locations = cells_row_groups(
      groups = everything()
    ),
    fn = function(x){ 
      map(x, \(y)
      case_match(
        y,
        "Plant~height~(cm)" ~ html("Plant height (cm)"),
        "Leaf~dry~mass~(g)" ~ html("Leaf dry mass (g)"),
        "Leaf~area~(cm^2)" ~ html("Leaf area (cm<sup>2</sup>)"),
        .default = html(y)
      )
      )
    }
  ) |> 
  tab_style(
        style = cell_borders(sides = "left", color = "white", weight = px(15), style = "solid"),
        locations = cells_body(
          columns = "Term")
      ) |>
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c("Term", 'Sum of Square', df, 'F', P),
      rows = P <= 0.05)
    )|>
  tab_options(
  table.font.size = 12,
  data_row.padding = gt::px(1)
  )

# table on the right
#kable(head(cars))
df |> 
  slice(31:60) |> 
  group_by(Traits) |> 
  gt() |> 
  text_transform(
    locations = cells_row_groups(
      groups = everything()
    ),
    fn = function(x){ 
      map(x, \(y)
      case_match(
        y,
        "Leaf~thickness~(mm)" ~ html("Leaf thickness (mm)"),
        "SLA~(cm^2*g^{-1})" ~ html("SLA (cm<sup>2</sup>g<sup>-1</sup>)"),
        "LDMC~(gg^{-1})" ~ html("LDMC (gg<sup>-1</sup>))"),
        .default = html(y)
      )
      )
    }
  ) |> 
  tab_style(
        style = cell_borders(sides = "left", color = "white", weight = px(15), style = "solid"),
        locations = cells_body(
          columns = "Term")
      ) |>
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c("Term", 'Sum of Square', df, 'F', P),
      rows = P <= 0.05)
    )|>
  tab_options(
  table.font.size = 12,
  data_row.padding = gt::px(1)
  )

```

```{r}
knitr::knit_exit()
```

### Multivariate analysis

```{r}
#| label: fig-trait_pca
#| echo: false
#| fig-height: 6
#| fig-cap: Multivariate traits

tar_read(trait_pca_figure)

```

## Supporting Information

### Trait correlations

The trait correlation is high between the size traits (area, height and dry mass).

```{r}
#| label: trait-correlation
#| echo: false
#| fig-cap: Correlations between traits.

tar_read(trait_correlation_plot)

```

### Trait coverage

The trait correlation is high between the size traits (area, height and dry mass).

```{r}
#| label: trait-n-imputation
#| echo: false
#| fig-width: 8
#| fig-cap: The proportions of the level traits were imputed for the warming x nitrogen model; plot (blue), block (green), site (orange) or global (red).

tar_read(imputation_plot_n)

```

```{r}
#| label: trait-g-imputation
#| echo: false
#| fig-cap: The proportions of the level traits were imputed for the warming x grazing model; plot (blue), block (green), site (orange) or global (red).

tar_read(imputation_plot_g)

```

### Species list

Species list at all the sites along the elevational gradient. The species are sorted by cover and functional group.

```{r}
#| label: tbl-sp-list
#| echo: false
#| tbl-cap: List of species at all three sites, sorted by cover and functional group.

tar_load(species_list_out)

tab <- species_list_out

tab %>% gtsave(filename = "Table_S2_species_list.png")

```

### PCA output

```{r}
#| label: tbl-pca
#| echo: false
#| eval: false
#| tbl-cap: PCA output.

tar_read(trait_pca)

pca_out <- trait_pca[[4]] |> 
  tidy() |> 
  mutate(term = case_when(str_detect(term, "warming:Nitrogen_log") ~ "WxN",
                            str_detect(term, "Nitrogen_log:origSiteID") ~ "NxO",
                            str_detect(term, "warming:origSiteID") ~ "WxO",
                            str_detect(term, "warming:grazing") ~ "WxG",
                            str_detect(term, "origSiteID:grazing") ~ "GxO",
                            str_detect(term, "origSiteID") ~ "O",
                            str_detect(term, "Nitrogen_log") ~ "N",
                            str_detect(term, "warming") ~ "W",
                            str_detect(term, "grazing") ~ "G",
                            TRUE ~ term),
         SumOfSqs = round(SumOfSqs, 2),
         R2 = round(R2, 2),
         statistic = round(statistic, 2)) |> 
  rename(Term = term, "Sum of squares" = SumOfSqs, Statistic = statistic, "P value" = p.value) |> 
  gt()

pca_out
pca_out %>% gtsave(filename = "Table_SI1.png")

```



### ITV output

```{r}
#| label: tbl-itv-origin
#| echo: false
#| tbl-cap: Proportion contribution ITV, turnover, ....

tar_load(variance_part)

dat <- variance_part |>
  select(origSiteID, figure_names, process, Term = term, pr = proportion, pv = pvalue) |>
  mutate(origSiteID = recode(origSiteID, "Alpine" = "A", "Sub-alpine" = "S")) |> 
  pivot_wider(names_from = c(origSiteID, process), values_from = c(pr, pv)) |> 
  select(figure_names, Term, pr_S_total, pv_S_total, pr_S_turnover, pv_S_turnover, pr_S_intraspecific, pv_S_intraspecific, pr_S_covariation, pr_A_total, pv_A_total, pr_A_turnover, pv_A_turnover, pr_A_intraspecific, pv_A_intraspecific, pr_A_covariation) |> 
  #select(trait_fancy, Term, pr_total, pv_total, pr_turnover, pv_turnover, pr_intraspecific, pv_intraspecific, pr_covariation) |> 
  group_by(figure_names)


tab <- dat |> 
  gt() |> 
  fmt_number(columns = everything(),
  decimals = 3) |> 
  text_transform(
    locations = cells_row_groups(
      groups = everything()
    ),
    fn = function(x){ 
      map(x, \(y)
      case_match(
        y,
        "Plant~height~(cm)" ~ html("Plant height (cm)"),
        "Leaf~dry~mass~(g)" ~ html("Leaf dry mass (g)"),
        "Leaf~area~(cm^2)" ~ html("Leaf area (cm<sup>2</sup>)"),
        "Leaf~thickness~(mm)" ~ html("Leaf thickness (mm)"),
        "SLA~(cm^2*g^{-1})" ~ html("SLA (cm<sup>2</sup>g<sup>-1</sup>)"),
        "LDMC~(gg^{-1})" ~ html("LDMC (gg<sup>-1</sup>))"),
        .default = html(y)
      )
      )
    }
  ) |> 
  tab_spanner(
    label = md("**Total**"),
    columns = c(pr_S_total, pv_S_total)
  ) |> 
  tab_spanner(
    label = md("**Turnover**"),
    columns = c(pr_S_turnover, pv_S_turnover)
  ) |> 
  tab_spanner(
    label = md("**ITV**"),
    columns = c(pr_S_intraspecific, pv_S_intraspecific)
  ) |> 
  tab_spanner(
    label = md("**Covariation**"),
    columns = c(pr_S_covariation)
  ) |> 
    tab_spanner(label = md("**Sub-Alpine**"), 
              columns = c(pr_S_total, pv_S_total, pr_S_turnover, pv_S_turnover, pr_S_intraspecific, pv_S_intraspecific, pr_S_covariation), 
              level = 2) |> 
  tab_spanner(
    label = md("**Total.**"),
    columns = c(pr_A_total, pv_A_total)
  ) |> 
  tab_spanner(
    label = md("**Turnover.**"),
    columns = c(pr_A_turnover, pv_A_turnover)
  ) |> 
  tab_spanner(
    label = md("**ITV.**"),
    columns = c(pr_A_intraspecific, pv_A_intraspecific)
  ) |> 
  tab_spanner(
    label = md("**Covariation.**"),
    columns = c(pr_A_covariation)
  ) |> 
    tab_spanner(label = md("**Alpine**"), 
              columns = c(pr_A_total, pv_A_total, pr_A_turnover, pv_A_turnover, pr_A_intraspecific, pv_A_intraspecific, pr_A_covariation), 
              level = 2) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c(Term, pr_S_total, pv_S_total),
      rows = pv_S_total <= 0.05)
    ) |> 
    tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c(Term, pr_S_turnover, pv_S_turnover),
      rows = pv_S_turnover <= 0.05)
    ) |> 
      tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c(pr_S_intraspecific, pv_S_intraspecific),
      rows = pv_S_intraspecific <= 0.05)
    ) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c(Term, pr_A_total, pv_A_total),
      rows = pv_A_total <= 0.05)
    ) |> 
    tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c(pr_A_turnover, pv_A_turnover),
      rows = pv_A_turnover <= 0.05)
    ) |> 
      tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c(pr_A_intraspecific, pv_A_intraspecific),
      rows = pv_A_intraspecific <= 0.05)
    ) |> 
  cols_label(
    Term = md("**Term**"),
    pr_S_total = md("Prop"),
    pv_S_total = md("*P*"),
    pr_S_turnover = md("Prop"),
    pv_S_turnover = md("*P*"),
    pr_S_intraspecific = md("Prop"),
    pv_S_intraspecific = md("*P*"),
    pr_S_covariation = md("Prop"),
    
    pr_A_total = md("Prop"),
    pv_A_total = md("*P*"),
    pr_A_turnover = md("Prop"),
    pv_A_turnover = md("*P*"),
    pr_A_intraspecific = md("Prop"),
    pv_A_intraspecific = md("*P*"),
    pr_A_covariation = md("Prop")
  ) |> 
  tab_options(
  table.font.size = 10,
  data_row.padding = gt::px(1)
  )

```

<!-- tab %>% gtsave(filename = "Table_2_itv_output.png") -->


```{r}


tab <- variance_part |>
  filter(process %in% c("intraspecific", "turnover")) |>
  select(origSiteID, figure_names, process, proportion) |> 
    group_by(origSiteID, Traits = figure_names, Process = process) |>
    summarise(proportion = sum(proportion)) |> 
  pivot_wider(names_from = origSiteID, values_from = proportion) |> 
  ungroup() |> 
  gt() |> 
  fmt_number(columns = everything(),
  decimals = 2) |> 
  cols_label(
    Traits = md("**Traits**"),
    Process = md("**Process**"),
    Alpine = md("**Alpine**"),
    'Sub-alpine' = md("**Sub-alpine**"),
  ) |> 
  tab_options(
  table.font.size = 10,
  data_row.padding = gt::px(1)
  )
#tab %>% gtsave(filename = "Table_3_itv_total.png")
```

