---
title: "Trait responses to global change"
format: html
editor: visual
execute:
  warnings: false
  messages: false
---

```{r}
#| label: set-up
#| echo: false

library(targets)
library(tarchetypes)
library(ggplot2)
library(ggcorrplot)
library(gt)
library(gtExtras)

```

## Methods

Run models without 3-way interaction and log-transform nitrogen addition.

### Warming x nitrogen addition

**log**: lm(trait \~ warming \* log(nitrogen) + warming \* origSiteID + log(nitrogen) \* origSiteID, data)

### Warming x grazing

**linear**: lm(trait \~ warming \* grazing + warming \* origSiteID + grazing \* origSiteID, data)

# Results

### Warming x Nitrogen

```{r}
#| label: tbl-single-traits
#| echo: false
#| fig-cap: Model output from single trait models

tar_load(n_trait_anova)

t <- n_trait_anova |> 
  mutate(sumsq = round(sumsq, 2),
         df = round(df, 2),
         statistic = round(statistic, 2),
         p.value = round(p.value, 3)) |> 
  rename(Traits = figure_names, Term = term, "Sum of Square" = sumsq, "F" = statistic, "P" = "p.value")

tab1 <- t |> 
  filter(Traits %in% c("Height~(cm)", "Dry~mass~(g)", "Area~(cm^2)")) |> 
  group_by(Traits) |> 
  gt() |> 
  tab_options(
  table.font.size = 12,
  data_row.padding = gt::px(1)
  ) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c(Term, "Sum of Square", df, "F", P),
      rows = P <= 0.05
    )
  ) |> 
  cols_align(
  align = c("left"),
  columns = Term
)
  
tab2 <- t |> 
  filter(!Traits %in% c("Height~(cm)", "Dry~mass~(g)", "Area~(cm^2)")) |> 
  group_by(Traits) |> 
  gt() |> 
  tab_options(
  table.font.size = 12,
  data_row.padding = gt::px(1)
  ) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c(Term, "Sum of Square", df, "F", P),
      rows = P <= 0.05
    )
  )  |> 
  cols_align(
  align = c("left"),
  columns = Term
)

listed_tables <- list(tab1, tab2)

gt_two_column_layout(listed_tables)
  
  

```

```{r}
#| label: fig-single-traits
#| echo: false
#| warnings: false
#| message: false
#| fig-height: 7
#| fig-cap: Trait responses to global change drivers

tar_read(n_trait_figure)

```

### Warming x Grazing

```{r}
#| label: fig-g-traits
#| echo: false
#| warnings: false
#| message: false
#| fig-height: 7
#| fig-cap: Trait responses to global change drivers

tar_read(g_trait_figure)

```

```{r}
#| label: tbl-wg-traits
#| echo: false
#| warnings: false
#| message: false
#| fig-cap: Model output from single trait models

tar_load(g_trait_anova)

g_trait_anova |> 
  ungroup() |> 
  select(-trait_trans) |> 
  mutate(sumsq = round(sumsq, 2),
         df = round(df, 2),
         statistic = round(statistic, 2),
         p.value = round(p.value, 3)) |> 
  rename(Traits = figure_names, Term = term, "Sum of Square" = sumsq, "F" = statistic, "P" = "p.value") |> 
  group_by(Traits) |> 
  gt() |> 
  tab_options(
  table.font.size = 12,
  data_row.padding = gt::px(1)
  ) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = c(Term, "Sum of Square", df, "F", P),
      rows = P <= 0.05
    )
  ) |> 
  cols_align(
  align = c("left"),
  columns = Term
)


```

### Multivariate analysis

```{r}
#| label: fig-trait_pca
#| echo: false
#| eval: false
#| fig-height: 10
#| fig-cap: Multivariate traits

tar_read(trait_pca_figure)

```

## Supporting Information

### Trait correlations

The trait correlation is high between the size traits (area, height and dry mass).

```{r}
#| label: trait-correlation
#| echo: false
#| eval: false
#| fig-cap: Correlations between traits.

tar_load(trait_correlation_plot)

trait_correlation_plot

```

### Trait coverage

The trait correlation is high between the size traits (area, height and dry mass).

```{r}
#| label: trait-imputation
#| echo: false
#| eval: false
#| fig-cap: The proportions of the level traits were imputed from before the bootstrapping procedure; plot (blue), block (green), site (orange) or global (red).

tar_load(imputation_plot)

imputation_plot

```
